---
title: "Homework 2 (p8105)"
author: "Britney Mazzetta (blm2156)"
output: github_document
---

# Problem 1

## Mr. Trash Wheel
```{r, eval = TRUE}
library(tidyverse)
library(readxl)

prob1 = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = 1, range = "A2:N338") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sport_balls = as.integer(sports_balls)) %>%
  select(-sports_balls)

```

## 2017 Precipitation
```{r}
precip2017 = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = 6, range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(
    Year = 2017
    )
```

## 2018 Precipitation
```{r}
precip2018 = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = 5, range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(
    Year = 2018
    )
```

## Combine Datasets

```{r}
combined_data =
  bind_rows(precip2017, precip2018) %>%
  mutate(
    month = as.numeric(month),
    month = month.name[month]
  )
  
```

## Paragraph

`r nrow(precip2017)`
`r nrow(precip2018)`
`r nrow(combined_data)`

`r median(pull(precip2017, total))`
`r median(pull(precip2018, total))`
`r median(pull(combined_data, total))`

The total amount of precipitation is `r sum(pull(precip2017, total))`
`r sum(pull(precip2018, total))`
`r sum(pull(combined_data, total))`

`r median(pull(prob1, sport_balls))`

# Problem 2

```{r}
pols = 
  read_csv("./data/pols-month.csv") %>%
  janitor::clean_names() %>% 
  separate(mon, c("year", "month", "day"), sep = "-")%>%
  mutate(
    year = as.numeric(year),
    month = as.numeric(month),
    month = month.abb[month],
    month = str_to_lower(month),
    president = ifelse(prez_gop == 1, "gop", "dem") 
  ) %>%
      select(-prez_gop, -prez_dem, -day)

```


```{r}
snp = 
  read_csv("./data/snp.csv") %>%
  janitor::clean_names() %>% 
  separate(date, c("month", "day", "year"), sep = "/")%>%
 select(year, month, -day, close)%>%
  mutate(
    month = as.numeric(month),
    month = month.abb[month],
    month = str_to_lower(month),
    year = as.numeric(year)
  ) %>%
  arrange(year, month)

  
```

```{r}
unemployment = 
  read_csv("./data/unemployment.csv") %>%
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment_rate"
  )
```
## Joining the datasets

```{r}
snp_pols = 
  left_join(pols, snp, by = c("year", "month"))

prob2_full =
  left_join(snp_pols, unemployment, by = c("year", "month"))

```
# Problem 3

```{r}
babies_data = read_csv("./data/Popular_Baby_Names.csv") %>%
  distinct() %>%
  janitor::clean_names() %>%
  mutate(ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER", "BLACK NON HISP" = "BLACK NON HISPANIC", "WHITE NON HISP" = "WHITE NON HISPANIC")) %>%
  mutate(ethnicity = str_to_lower(ethnicity), gender = str_to_lower(gender), childs_first_name = str_to_lower(childs_first_name))
```

```{r, include = TRUE}
babies_data %>%
filter(childs_first_name == "olivia") %>%
  select(-count) %>%
  pivot_wider(names_from = year_of_birth,
      values_from = rank)%>%
  knitr::kable()
```
```{r}
babies_data %>%
filter(gender == "male", rank == 1) %>%
  select(-count) %>%
  pivot_wider(names_from = year_of_birth,
      values_from = rank) %>%
 knitr::kable()
```
```{r}
babies_plot = filter(babies_data, ethnicity == "white non hispanic", year_of_birth == "2016", gender == "male")
ggplot(babies_plot, aes(x = rank, y = count)) + geom_point()
```




