---
title: "Homework 2 (p8105)"
author: "Britney Mazzetta (blm2156)"
output: github_document
---

# Problem 1

## Mr. Trash Wheel
```{r, eval = TRUE}
library(tidyverse)
library(readxl)

prob1 = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = 1, range = "A2:N338") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sport_balls = as.integer(sports_balls)) %>%
  select(-sports_balls)

```

## 2017 Precipitation
```{r}
precip2017 = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = 6, range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(
    Year = 2017
    )
```

## 2018 Precipitation
```{r}
precip2018 = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = 5, range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(
    Year = 2018
    )
```

## Combine Datasets

```{r}
combined_data =
  bind_rows(precip2017, precip2018) %>%
  mutate(
    month = as.numeric(month),
    month = month.name[month]
  )

```



## Problem 1 Summary

For this dataset, we are looking at the Mr. Trash Wheel dataset, in addition to preciptation data for 2017 and 2018. The Mr. Trash Wheel dataset (prob 1), contains several variables pertaining to specific dumpsters. Some of these variables include the year and date information on that dumpster was collected, the weight in tons and volume of the trash collected in each dumpster, in addition to different types and their corresponding amounts of trash collected (plastic bottles, polystyrene, cigarette butts, glass bottles, grocery bags, chip bags, homes powered, sport balls). The average number of cigarette butts in a dumpster across the whole dataset was `r mean(pull(prob1, cigarette_butts))`. The median number of sports balls in a dumpster in 2017 was 8. 
```{r}
prob1 %>% 
  filter(year == 2017) %>% 
  summarise(median_balls = median(sport_balls))
```
The 2017 Precipiation dataset (precip2017) contains the amount of rainfall per month in 2017. The median amount of rain in 2017 was `r median(pull(precip2017, total))` while the mean amount of rain in 2017 was `r mean(pull(precip2017, total))`. The total amount of precipitation is in 2017 was `r sum(pull(precip2017, total))`. 

The 2018 Precipiation dataset (precip2018) contains the amount of rainfall per month in 2018. The median amount of rain in 2018 was `r median(pull(precip2018, total))` while the mean amount of rain in 2018 was `r mean(pull(precip2018, total))`. The total amount of precipitation is in 2018 was `r sum(pull(precip2018, total))`. 

The combined dataset (combined_data) contains the data collected from both the 2017 Precipitation dataset and the 2018 Precipitation dataset. There are now `r nrow(combined_data)` rows and `r ncol(combined_data)` columns in the combined dataset. The median amount of rain in the combined dataset (for years 2017 and 2018) was `r median(pull(combined_data, total))` while the mean amount of rain in the combined dataset (for years 2017 and 2018) was `r mean(pull(combined_data, total))`. The total amount of precipitation in the combined dataset (for years 2017 and 2018) was `r sum(pull(precip2017, total))`. 


# Problem 2

## Pols Data 
```{r}
pols = 
  read_csv("./data/pols-month.csv") %>%
  janitor::clean_names() %>% 
  separate(mon, c("year", "month", "day"), sep = "-")%>%
  mutate(
    year = as.numeric(year),
    month = as.numeric(month),
    month = month.abb[month],
    month = str_to_lower(month),
    president = ifelse(prez_gop == 1, "gop", "dem") 
  ) %>%
      select(-prez_gop, -prez_dem, -day)

```

## SNP Data
```{r}
snp = 
  read_csv("./data/snp.csv") %>%
  janitor::clean_names() %>% 
  separate(date, c("month", "day", "year"), sep = "/")%>%
 select(year, month, -day, close)%>%
  mutate(
    month = as.numeric(month),
    month = month.abb[month],
    month = str_to_lower(month),
    year = as.numeric(year)
  ) %>%
  arrange(year, month)

  
```

## Unemployment Data
```{r}
unemployment = 
  read_csv("./data/unemployment.csv") %>%
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment_rate"
  )
```

## Joining the datasets

```{r}
snp_pols = 
  left_join(pols, snp, by = c("year", "month"))

prob2_full =
  left_join(snp_pols, unemployment, by = c("year", "month"))

```

## Problem 2 Summary 
This problem utilized the FiveThirtyEight data. Through this, we looked at several different datasets: pols data, snp data, unemployment data, and a combination of data. 

The pols data contains `r nrow(pols)` observations of `r ncol(pols)` variables. These `r ncol(pols)` variables indicates whether the president was republican on the associated date, indicates the number of republican governors, senators, and representatives on the associated date, in addition to the number of democratic governors, senators, and representatives on the associated date. The years for this dataset spanned from 1947 to 2015.

The snp data contains `r nrow(snp)` observations of `r ncol(snp)` variables related to Standard & Poor's stock market index. These `r ncol(snp)` variables described the stock market date observed and the closing values of the S&P index on the associated date. The years for this dataset spanned from 1950 to 2015.

The unemployment data contains `r nrow(unemployment)` observations of `r ncol(unemployment)` variables. The variables described the percentage of unemployment for each of the 12 months in the associated year. The years for this dataset spanned from 1948 to 2015.

The combined dataset (snp_pols) combined all of the information from both the pols and snp datasets respectively for the appropriate corresponding dates. This new dataset had the same number, `r nrow(snp_pols)`, observations and an additional variable with `r ncol(snp_pols)` total variables.

# Problem 3

```{r}
babies_data = read_csv("./data/Popular_Baby_Names.csv") %>%
  distinct() %>%
  janitor::clean_names() %>%
  mutate(ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER", "BLACK NON HISP" = "BLACK NON HISPANIC", "WHITE NON HISP" = "WHITE NON HISPANIC")) %>%
  mutate(ethnicity = str_to_lower(ethnicity), gender = str_to_lower(gender), childs_first_name = str_to_lower(childs_first_name))
```

```{r, include = TRUE}
babies_data %>%
filter(childs_first_name == "olivia") %>%
  select(-count) %>%
  pivot_wider(names_from = year_of_birth,
      values_from = rank)%>%
  knitr::kable()
```
```{r}
babies_data %>%
filter(gender == "male", rank == 1) %>%
  select(-count) %>%
  pivot_wider(names_from = year_of_birth,
      values_from = rank) %>%
 knitr::kable()
```
```{r}
babies_plot = filter(babies_data, ethnicity == "white non hispanic", year_of_birth == "2016", gender == "male")
ggplot(babies_plot, aes(x = rank, y = count)) + geom_point()
```




